Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2017-11-09 TACAS2018/paper.tex
Optimization algorithm

Our abstraction refinement CEGIS loop is similar to CEGIS with counterexample generalization~\cite{david2017program}.
However, it provides several improvements by exploiting modularity inside the loop.
The novelty of our approach relies on the synthesis refinement phase, which enables us
to generalise an existing candidate by means of optimisation. Rather than allow unsuitable
candidates reach the verification phase, and possibly require further induction on the synthesiser,
we use an optimizing algorithm which finds a local minima for the properties of interest. 
This reduces the chances of the optimised candidate failing verification. Furthermore, it allows
us to relax the quality of the synthesis, since unlikely candidates will still be optimised during this phase. 
Optimization may also enter its own refinement loop, since it is possible for the
counterexample to indicate which properties need optimizing (i.e. the ones that are failing).

\todo{I assume the optimization problem is something like: ``find a controller and an observer that minimize some cost function''. What is the cost function in this case? I guess you want something like the
  difference between the current candidate and the general solution? How would we capture that?}

\todo{What is the optimization alg that you use? Point 4 below doesn't specify much.}

@@

\item In {\sc synthesize}, we synthesize a candidate controller $\mat{K}$
and observer $\mat{L}$ that satisfy
$\phi_{synth}$ by invoking a SAT solver.
There are two exit conditions:
If there is no candidate solution we return FAIL and signal {\sc Abstract}. Based on the current abstraction,
{\sc  Abstract} may choose to increase the precision, refine the abstraction, or return a synthesis failure.
If we have a candidate solution we will feed it to the verifier chain in order to validate it.
%
\item {\sc Verify Synthesis} verifies $\phi_{static}$ of  both the
controller and the observer.  If one of these results does not satisfy the tighter
specification, it feeds back to the {\sc Refine Synthesis} phase which uses sensitivity analysis to
synthesise a refined solution. If the candidate is valid, we feed it to {\sc Discrete Time}. 
\item {\sc Refine Synthesis} takes $\mat{K},\mat{L} \not \models \phi_{static}$ and synthesises $\mat{K}_{\langle C \rangle} = \mat{K}+\mat{\delta}_K, \mat{L}_{\langle C \rangle}=\mat{L}+\mat{\delta}_L, where \mat{K}_{\langle C \rangle}, \mat{L}_{\langle C \rangle} \models \phi_{static}$
If we cannot synthesise a suitable candidate, we signal the {\sc Abstract} phase to refine $\phi_{synth}$ and request {\sc Synthesise} to produce a new candidate. Otherwise the refined candidate solution is given back to {\sc Verify Synthesis}.
%
\item {\sc Discrete Time} verifies $\phi_{safety}^{\infty}$ using discrete
time dynamics.  If the verification fails, we report the counterexample to
{\sc Refine Abstraction} including details of initial states and
iterations that failed the specification.  
\item {\sc Refine Abstraction} changes $\phi_{safety}$ to encompass any arising conterexample. 
This can be done by increasing the time horizon $k$ for the safety specification of the violating initial condition and input.
It also modifies $\phi_{static}$ to account for the counterexample.
If the magnitude of the violation is small, {\sc Refine Synthesis} may find a new candidate, otherwise it will feed through to {\sc Abstract}.
%
\item {\sc Continuous Time} verifies that the
transitions between sample times meet the specification $\phi_{safety}^{\infty}$.  If
a counterexample is found, information about the maximum distance of the
violation to the specification boundary is fed to {\sc Sample}. 
\item {\sc Sample} will select a smaller suitable sample time and modify the abstractions to meet the discretization. 
It will then feed back to the abstraction refinement loop to look for a suitable action for the updated model.
%

@@

>>>>

Our abstraction refinement CEGIS loop is similar to CEGIS with counterexample generalization~\cite{david2017program}.
However, it provides several improvements by exploiting modularity inside the loop.
The novelty of our approach relies on the synthesis refinement phase, which
enables us to generalise an existing candidate by means of optimisation. Rather
than allow unsuitable candidates reach the verification phase, and possibly
require further induction on the synthesiser, we use an optimizing algorithm
which finds a local minima for the properties of interest.  This reduces the
chances of the optimised candidate failing verification.  Furthermore, it
allows us to relax the quality of the synthesis, since unlikely candidates will
still be optimised during this phase. 
The optimization is performed on an objective function which typically depends
on a subset of the properties.  Hence the chosen optimization function depends
on which properties fail verification. Optimization may also enter its own
refinement loop, since it is possible for the counterexample to indicate which
properties need optimizing (i.e. the ones that are failing).
In our control synthesis problem, verification failures are typically due to
safety (since stability and response time are maintained in the highest
abstraction levels).  The optimization function therefore looks to minimize
the reachspace by making the dynamics as convergent as possible (note that
this also improves the response time of the system).

@@

\item In {\sc synthesize}, we synthesize a candidate controller $\mat{K}$
and observer $\mat{L}$ that satisfy
$\phi_{synth}$ by invoking a SAT solver.
There are two exit conditions:
If there is no candidate solution we return FAIL and signal {\sc Abstract}. Based on the current abstraction,
{\sc  Abstract} may choose to increase the precision, refine the abstraction, or return a synthesis failure.
If we have a candidate solution we will feed it to the verifier chain in order to validate it.
%
\item {\sc Verify Synthesis} verifies $\phi_{static}$ of  both the
controller and the observer.  If one of these results does not satisfy the tighter
specification, it feeds back to the {\sc Refine Synthesis} phase which uses sensitivity analysis to
synthesise a refined solution. If the candidate is valid, we feed it to {\sc Discrete Time}. 
\item {\sc Refine Synthesis} takes $\mat{K},\mat{L} \not \models \phi_{static}$ and synthesises $\mat{K}_{\langle C \rangle} = \mat{K}+\mat{\delta}_K, \mat{L}_{\langle C \rangle}=\mat{L}+\mat{\delta}_L, where \mat{K}_{\langle C \rangle}, \mat{L}_{\langle C \rangle} \models \phi_{static}$.
Furthermore we perform optimization based on the violation of the safe-space (if found).
Given a closed loop controller or observer block ($\mat{A}_c$ or $\mat{A}_o$),
we look for a constant $c \leq 1$ such that $\mat{B}\mat{K}'=(1-c)\mat{A}_d+c\mat{B}\mat{K} \mid \mat{K}' \models \phi_K$
(and similarly for $\mat{L}'$.
If we cannot synthesise a suitable candidate, we signal the {\sc Abstract} phase to refine $\phi_{synth}$ and request {\sc Synthesise} to produce a new candidate. Otherwise the refined candidate solution is given back to {\sc Verify Synthesis}.
%
\item {\sc Discrete Time} verifies $\phi_{safety}^{\infty}$ using discrete
time dynamics.  If the verification fails, we report the counterexample to
{\sc Refine Abstraction} including details of initial states and
iterations that failed the specification.  
\item {\sc Refine Abstraction} changes $\phi_{safety}$ to encompass any arising conterexample. 
This can be done by increasing the time horizon $k$ for the safety specification of the violating initial condition and input.
It also modifies $\phi_{static}$ to account for the counterexample.
If the magnitude of the violation is small, {\sc Refine Synthesis} may find a new candidate, otherwise it will feed through to {\sc Abstract}.
%
\item {\sc Continuous Time} verifies that the
transitions between sample times meet the specification $\phi_{safety}^{\infty}$.  If
a counterexample is found, information about the maximum distance of the
violation to the specification boundary is fed to {\sc Sample}. 
\item {\sc Sample} will select a smaller suitable sample time and modify the abstractions to meet the discretization. 
It will then feed back to the abstraction refinement loop to look for a suitable action for the updated model.
%

@@

Our Synthesis Algorithm shows promise in exploiting modularity for
improved response. It however relies on the use of application
specific optimization which means that it is necessary to find adequate tools for
exploring different types of properties.  Our results show significant
improvement in terms of scalability over standard approaches with an ability to
explore complex properties, hence many synthesis tasks may be derived
from it. In particular we note that our synthesis loop should be targeted at
systems where the solution space is formed by connected spaces where
optimization algorithms or localized synthesis are likely to be achieved more
efficiently than global synthesis.
Finally our application to control synthesis provides a solid proof
of concept for scalable optimal control synthesis of LTI systems. There is still
a need to extend it it MIMO systems and analyse further properties, which may
also require exploring other optimization methods for control systems.
